---
title: "Salmon"
author: "Gracie White"
date: "November 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part 1 - Load packages and Read Data
```{r, echo=FALSE}
library(tidyverse)

salmon <- read_csv("salmon_imports copy.csv")

```
#Part 2. Exploratory Scatterplot

```{r}
salmon_plot <- ggplot(salmon, aes(x = year, y = salmon_fresh)) + 
  geom_point()
salmon_plot


##Overall, this looks fairly linear apart a few instances. I have concerns about extrapolting more than a few years into the future, and would like to know if there are certain events that led to the variations in 2004 and 2014?
```
#Part 3 - linear regression 

```{r}
salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model

#Intercept: -14,982,940
#y=-14,982,940 + 7550x + Error

# Imports(ton) = 7750(year)+ -14,982,940

#Every year, salm imports increase by approximately 7550 tons. (slope)
#At year zero, imports would = -14,982,940, which is impossible, so we can't extrapolate that far.



```
#part 4 - Model Diagnostics 

```{r}
plot(salmon_model)
#residuals appear mostly normal, but maybe with a slight increase in recent years (showin the residuals vs fitted plot). QQ plot shows a couple outliers. 

par(mfrow = c(2,2))
plot(salmon_model)
```
#Part 5 - Model fit and significance

```{r}
summary(salmon_model)

#year does significantly predict salmon imports within the scope of our model. Imports can be predicted 95% of the time by our model. 5% of our data is not explained by our model. 
```

#Part 6 - a final graph!

```{r}
salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point() +
 geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") +
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) + 
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")
salmon_final_graph
  
```
#Part 7 - Predictions!

```{r}
new_years <- data.frame(year = c(2022, 2024, 2026, 2028))
future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")
future_predict

predictions <- data.frame(new_years, future_predict)
predictions


```
#Part 8 - Pearson's r

```{r}
salmon_cortest <- cor.test(salmon$year, salmon$salmon_fresh)
salmon_cortest

#Pearson's r = 0.973
#A strong correlation

```
#The model (Imports(ton) = 7750(year)+ -14,982,940) explains a significant amount of variance in fresh salmon imports (*p* < 0.0001, R^2^ = 0.95, Pearson's *r* = 0.97). There is a storng postive crrelation between the variables, although caution should be used when extrapolating. 
